An Extendable \proglang{Python} Implementation of Robust Optimisation Monte Carlo

### 2022-02-20 article/article.tex
final changes

\newcommand{\vb}{\mathbf{v}}
\newcommand{\ub}{\mathbf{u}}
\newcommand{\yb}{\mathbf{y}}
\newcommand{\xb}{\mathbf{x}}

@@

%% - \Abstract{} almost as usual
\Abstract{Performing inference in statistical models with intractable
  likelihood is challenging; most likelihood-free inference (LFI)
  methods encounter accuracy and efficiency limitations. In this
  paper, we present the implementation of Robust Optimisation Monte
  Carlo (ROMC) in the \proglang{Python} package \pkg{ELFI}. ROMC is a
  novel and efficient (highly-parallelisable) LFI framework that
  provides accurate weighted samples from the posterior. Our
  implementation can be used in two ways. First, a scientist may use
  it as an out-of-the-box LFI algorithm; we provide an easy-to-use API
  harmonised with the principles of \pkg{ELFI}, enabling effortless
  comparisons with the rest of the methods included in the
  package. Additionally, we have carefully split ROMC into isolated
  components for supporting extensibility. A researcher may experiment
  with novel method(s) for solving part(s) of ROMC without
  reimplementing everything from scratch. In both scenarios, all steps
  run in a fully-parallelisable manner, exploiting all CPU
  cores. We also provide helpful functionalities for (i) inspecting
  the inference process and (ii) evaluating the obtained
  samples. Finally, we test the robustness of our implementation on
  some typical LFI examples.}

@@

%% -- Design Principles -----------------------------------------------------
\section{Implementation}
This section is split in two parts. We first express ROMC as an
algorithm and then we present the general implementation principles we
follow.

\subsection{Algotirhmic view of ROMC}
\input{latex_files/ch3/3-0-design-principles.tex}

@@

\section*{Computational details}
The results in this paper were obtained using \proglang{Python}~3.7.9
with the \pkg{ELFI}~0.8.3 package. The experiments have been executed
in a single machine with an Intel® Core™ i7-8750H Processor an with
Ubuntu 20.04 lts operating system.

@@

HP was funded by European Research Council grant 742158 (SCARABEE,
Scalable inference algorithms for Bayesian evolutionary epidemiology).

>>>>

\newcommand{\vb}{\mathbf{v}}
\newcommand{\yb}{\mathbf{y}}
\newcommand{\xb}{\mathbf{x}}

@@

%% - \Abstract{} almost as usual
\Abstract{Performing inference in statistical models with intractable
  likelihood is challenging; most likelihood-free inference (LFI)
  methods encounter accuracy and efficiency limitations. In this
  paper, we present the implementation of Robust Optimisation Monte
  Carlo (ROMC) in the \proglang{Python} package \pkg{ELFI}. ROMC is a
  novel and efficient (highly-parallelisable) LFI framework that
  provides accurate weighted samples from the posterior. Our
  implementation can be used in two ways. Firstly, a scientist may use
  it as an out-of-the-box LFI algorithm; we provide an easy-to-use API
  harmonised with the rest of the \pkg{ELFI} framework for enabling
  effortless comparisons with the rest of the LFI methods implemented
  there. Additionally, we support extensibility by splitting ROMC into
  isolated components. A researcher may experiment with novel
  method(s) for solving specific subpart(s) of ROMC without
  reimplementing everything from scratch. In both scenarios, the
  algorithm runs in a fully-parallelisable manner for exploiting all
  CPU cores. We also provide helpful functionalities for (i)
  inspecting the inference process and (ii) evaluating the obtained
  samples. Finally, we test the robustness of our implementation on
  some typical LFI examples.  }

@@

%% -- Design Principles -----------------------------------------------------
\section{Design principles}  % better title needed
\input{latex_files/ch3/3-0-design-principles.tex}

@@

\section*{Computational details}

The results in this paper were obtained using \proglang{Python}~3.7.9
with the \pkg{ELFI}~0.8.3 package. The experiments have been executed
in a single machine with an Intel® Core™ i7-8750H Processor an with
Ubuntu 20.04 lts operating system.

@@

Henri Pesonen was funded by European Research Council grant 742158 (SCARABEE,
Scalable inference algorithms for Bayesian evolutionary epidemiology).
